\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-25 paper/paper.tex
Edits and a couple fixes to plotting functions.

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

@@

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or where in operation.  The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.  Figure \ref{fig1} shows a scatterplot of the total observed running time in hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analyses.

@@

\subsection{Single GLFP Model}
\label{subsec:ex1}
We fit the GLFP model on real hard drive failure data from a cloud-based storage company.  Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model.  The data contain the number of hours the hard drive was in operation, and whether the drive failed or was right censored.  No information is available about the cause of failure.  To illustate the GLFP model, we will present an analysis of the drive-model with the most observed failures. We use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}$) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}$). Although non-standard, for purposes of exposition, we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.

%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, 5.6\times 10^7 \rangle$$
$$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
\sigma_1 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle, \quad
\sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.  Given the interpretation of $\pi$ as the proportion of defective units, which we would expect to be small, a range primarily covering the left side of 0.50 is reasonable and weakly informative.  The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.  Again, this allows for a wide range of hazards, larger than we would expect to see in the data.  The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$. For $t_{0.2,2}$ the range is $(3.4,16.6)$. Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.  They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower mean time to failure than the wearout mode.  This is helpful for estimation since, as discussed in Chan \& Meeker, ML had difficulty identifying the two sets of GLFP parameters without making strong assumptions about the data.  \\

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.  The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter estimates for the two Weibull distributions also match intuition.  The wearout failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, an increasing hazard.  The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
$\pi$ & 0.03 & 0.05 & 0.10 \\ 
 $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
\label{table:1}
\end{table}

@@

In Figure 1 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.  The solid stepwise curve corresponds to the observed failures.  Censored drives are not plotted.  Standard error bands are calculated using Greenwood's method \cite{green}.  Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.  The non-parametric and parametric model align up quite closely.  Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.  Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution. The discrepancy in the right tail of the distribution appears to be real. We don't provide a solution to this problem here. Certainly, if there are clearly more that two well-separated failure modes, the GLFP will not fit the data well. However, we consider it satisfactory if we can provide a good approximation to the observed lifetime distribution over the bulk of the product life. In practical terms, the right tail is of less importance for electronic goods, due to obsolescence.

>>>>

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

@@

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or where in operation.  The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.

@@

\subsection{Single GLFP Model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures (1,713), drive-model 14.  The non-parametric Kaplan-Meier estimate plotted below does not appear linear, suggesting different rates of failure in the data. In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.  Thus, as a first step, the GLFP model seems like an appropriate parametric model.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{KM14.pdf}
  \caption{Kaplan Meier estimate for drive model 14.  Bands correspond to Greenwood standard errors.}
  \label{fig1}
\end{figure}

To estimate the GLFP we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}$) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}$). Although non-standard, for purposes of exposition, we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.

%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, 5.6\times 10^7 \rangle$$
$$\mbox{logit}^{-1}\pi \sim \op{Logitnormal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
\sigma_1, \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle 
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.  Given the interpretation of $\pi$ as the proportion of defective units, which we would expect to be small, a range primarily covering the left side of 0.50 is reasonable and weakly informative.  The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.  Again, this allows for a wide range of hazards, larger than we would expect to see in the data.  The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$. For $t_{0.2,2}$ the range is $(3.4,16.6)$. Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative. \\

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.  The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter estimates for the two Weibull distributions also match intuition.  The wearout failure mode puts posterior probability on a values of $\sigma_1$ greater than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\sigma_2$, the wearout mode, is strictly below 1, an increasing hazard.  The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
$\pi$ & 0.03 & 0.05 & 0.10 \\ 
 $\sigma_1$ & 1.25 & 1.96 & 15.54 \\ 
  $\sigma_2$ & 0.48 & 0.65 & 0.78 \\ 
  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
\label{table:1}
\end{table}

@@

In Figure 3 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.  The solid stepwise curve corresponds to the observed failures.  Censored drives are not plotted.  Standard error bands are again calculated using Greenwood's method \cite{green}.  Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.  The non-parametric and parametric model align up quite closely.  Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.  Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution. The discrepancy in the right tail of the distribution appears to be real. We don't provide a solution to this problem here. Certainly, if there are clearly more that two well-separated failure modes, the GLFP will not fit the data well. However, we consider it satisfactory if we can provide a good approximation to the observed lifetime distribution over the bulk of the product life. In practical terms, the right tail is of less importance for electronic goods, due to obsolescence.
